---
title: "Homework Assignment 6"
author: "Arman Aghamyan"
date: "14/04/2019 "
output: html_document
---

You are required to submit both Markdown and HTML files with your answers and code in it. Do not remove problems from your Markdown file. Please write your name at the beginning of your file into the YAML-header inside of author. 

Data set relevant for this assignment can be found on Moodle. The description of the variables is given in a separate file. 

Good luck!

```
                                               Task 1   (4 pt.)
a. Clean and describe the final data. Which of variables can be useful for survival analysis? Why? Which of them can be an event indicator?
b. What is the difference between interval censored and point censored data?
c. What are the mail causes of censored data (right-censored)? Why  cannot we ignore censoring, bring and example?

```
```{r}
library(dplyr)
library(tidyr)
set.seed(1)
surv<-read.csv("survdata.csv")
str(surv)
surv<-na.omit(surv)
any(duplicated(surv$X))
surv<-select(surv,-X)
surv$alcohol<-ifelse(surv$alcohol>0,1,0)
surv$smoke<-ifelse(surv$smoke>0,1,0)
surv$hospital<-ifelse(surv$hospital==1,1,0)
surv$alcohol<-as.factor(surv$alcohol)
surv$smoke<-as.factor(surv$smoke)
surv$region<-as.factor(surv$region)
surv$poverty<-as.factor(surv$poverty)
surv$race<-as.factor(surv$race)
surv$bweight<-as.factor(surv$bweight)
glimpse(surv)

```
A)Now we have cleaned data with 9 categorical and 7 numeric variables.All variables can be usefull,cause they describe what can be fact that make child ill and the indicator event will be hospital("1" as child is ill or hospitlized).

B)Interval censored meanes that we know in which interval we have censored data,but exactly when we can not say.
Point censored data exactly mention when we have censored our data,we can say when we lost our datapoint.

C)For example, let the time-to-event be a person's age at onset of cancer. If you stop following someone after age 65, you may know that the person did NOT have cancer at age 65, but you do not have any information after that age.

You know that their age of getting cancer is greater than 65. But you do not know if they will never get cancer or if they'll get it at age 66, only that they have a "survival" time greater than 65 years. They are censored because we did not gather information on that subject after age 65.

So one cause of censoring is merely that we can't follow people forever.  At some point you have to end your study, and not all people will have experienced the event.

But another common cause is that people are lost to follow-up during a study. This is called random censoring. It occurs when follow-up ends for reasons that are not under control of the investigator.

In survival analysis, censored observations contribute to the total number at risk up to the time that they ceased to be followed.  One advantage here is that the length of time that an individual is followed does not have to be equal for everyone. All observations could have different amounts of follow-up time, and the analysis can take that into account.

```
                                               Task 2   (6 pt.)
  Create survival object to solve the task, use the KM method.
  
  a. What is the probability of surviving to a certain point in time (baseline model without predictors)? Use n.risk and n.event of summary function to comment on this task. 
  b. Why is the KM model called non-parametric?
  c. Plot the survival probability and cumulative hazard functions. Comment on it.
  d. Add independent variable/s (with more than 2 categories) to your model to create survival curves for different groups. 
  e. Find the variable with significant differences in survival probabilities. Show the survival plot and add a risk table to the graph. Comment on it.
                                               
                                               
```
A)The probability that child will not get to the hospital after month 3 is 0.990. n.risk is 3105,n.event is 12.

B)Non-parametric (or distribution-free) inferential statistical methods are mathematical procedures for statistical hypothesis testing which, unlike parametric statistics, make no assumptions about the probability distributions of the variables being assessed.One of them is Kaplan-Meier.

C)From surviving plot we see the difference very small aproximately 2,5%,hazard is the probability of get in to the hospital(not surviving) and in this plot it is more visually seen how much by each month the probability to get hospitalizied (not to survive) is increaseing.

E)As we see children with birthweight < 5.5 aproximately by 80% are risky to get hospiatlized than children with high birth wieght.
```{r,fig.width=8,fig.height=6}
library(survival)
library(survminer)

# D)
set.seed(1)
surv_object<-Surv(time = surv$chldage,event = surv$hospital)
km_1<-survfit(surv_object~1,data=surv)
summary(km_1)
table(surv$chldage,surv$hospital)
ggsurvplot(km_1,data=surv) # survive
ggsurvplot(km_1,fun="cumhaz",conf.int = F) #hazard
km_ra<-survfit(surv_object~race+alcohol,data=surv)
summary(km_ra)
ggsurvplot(km_ra,conf.int = F,ylim=c(0.9,1))
survdiff(surv_object~bweight,data=surv)
km_reg<-survfit(surv_object~bweight,data=surv)
ggsurvplot(km_reg,conf.int = F,ylim=c(0.9,1),risk.table = T)

```


```
                                               Task 3   (4 pt.)
a. Run the ATF model with independent variables (at least one numeric and one categorical). Overall model and coefficients must be significant. Interpret the coefficients.
b. Make predictions based on your model using quantiles. Why do we not use the predict function?
c. Plot probability of a hazard and survival. Comment on it.


```
A)One unit increase in education year increases survival time by 20%.
  When mother's race is 2nd category the survival time is increased by 21%.
  
B)As we want have predictions in time interval we should use quantile method,as response method brings the mean or median,and it is not effective to have such predictions.It does not give us usefull information.

C)As we see by plot by time increase the probability of survival decrease,but probability of hazard increases.
```{r}
set.seed(1)
reg_model<-survreg(surv_object~education+race,data=surv,dist = "exponential")
summary(reg_model)
exp(coef(reg_model))
quant=seq(.1,.9,length=9)
pred<-predict(reg_model,type="quantile",p=quant)
head(pred,n=10)
time<-1:12
pred_p<-pexp(time,rate=1/pred[1])
pred_p[1:10]

pred_p_surv=1-pred_p
ggplot()+geom_line(aes(x=time,y=pred_p_surv))+labs(x="Time",y="Probability of survival")

pred_p_hazard=-log(pred_p_surv)
ggplot()+geom_line(aes(x=time,y=pred_p_hazard))+labs(x="Time",y="Probability of hazard")

```

```
                                               Task 4   (6 pt.)

a. Which are the key assumptions of the Cox model?
b. Run the Cox proportional hazard model with significant coefficients. What does Rsquare show? 
c. Interpret the coefficients of your model. What does HR indicate?
d. Check the proportional hazard assumption. Comment on it.
e. Make predictions with cox model using plots.

```
A)The basic Cox model assumes that the hazard functions for two different levels of a covariate are proportional for all values of t.
Hazard ratio is independent from time.The hazard ratio is the same for all time periods.

B)Here R_squared is called a pseudo R_squared which shoes difference compared with basline model(without any predictor).The limit of pseudo R_squared is not 1.The max_possible R_squared shows maximum possible R-squared that we can get with perfect model.Here we have pseudo R_squared=0.006 which shoes that we have not a high difference compared with baseline model.And maximum possible R_squared is 0.289.

C)HR<1.Education reduce hazard,thus increased survival time.One unit increase in education year decreases probability of hazard by 16%.
  HR>1.Smoking increases hazard,thus decreases survival time.Smoking increases hazard by 100%.
  
D)  H0: Hazard rates are proportional
   
     H1: Hazard rates are not proportional
   
  As we see p_value is >0.05 so we fail to reject H0:hypothesis 
  So rates are proporional.

```{r}
library(ggplot2)
set.seed(1)
cox_model<-coxph(surv_object~education+smoke,data=surv)
summary(cox_model)
exp(coef(cox_model))
test_z<-cox.zph(cox_model,global=T,transform = "rank")
test_z

# E)
cox.zph(cox_model)
case<-data.frame(smoke="1",education=7)
case_surv<-survfit(cox_model,newdata = case)
names(case_surv)
data<-data.frame(Probability=case_surv$surv,Time=case_surv$time)
ggplot(data,aes(x=Time,y=Probability))+geom_point()+ggtitle("Survival Probabilities") #survive

data_1<-data.frame(Probability=case_surv$cumhaz,Time=case_surv$time)
ggplot(data_1,aes(x=Time,y=Probability)) + geom_point()+ ggtitle("Survival Probabilities") +
  xlab("Time") + ylab("Cumulative Hazard")

```